DESING AND SIMULATION SOFTWARE FOR RHODOTRON TYPE ELECTRON ACCELERATORS

### 2024-03-23 subsections/design/magnet/magnet.tex
hard copy ready

In rhodotron accelerators, magnets are used for not only guiding the electron beam to desired path, but also for synchronizing them with the RF field in the cavity. 
This syncronization, also called phase stability, is the fundamental constraint of magnet design. 
Time spent outside of the RF cavity should be precisely tuned to maximize acceleration in the succeeding pass. 
\iffalse \begin{figure}[H]
    %\captionsetup[subfigure]{justification=centering}
    %\captionsetup{justification=centering}
    \centering
    \begin{subfigure}{.5\textwidth}
      \centering
      \includegraphics[width=.9\linewidth]{../../../figures/design/Ltot.png}
    \end{subfigure}%
    \centering
    \begin{subfigure}{.5\textwidth}
      \centering
      \includegraphics[width=.9\linewidth]{../../../figures/design/tt300_mag.png}
    \end{subfigure}
    \caption{Geometry and modeling, outside trajectory of a rhodotron (TT300) \cite{cite:rhodo_design}.}
    \label{fig:magnet_design_illustrations}
\end{figure} \fi
\vspace{20pt}
\begin{figure}[H]
    \centering
    \subfigure{{\includegraphics[width=.45\textwidth]{../../../figures/design/Ltot.png} }}%
    \qquad\subfigure{{\includegraphics[width=.45\textwidth]{../../../figures/design/tt300_mag.png} }}%
    \vspace{0pt}
    \caption{\centering Geometry and modeling, outside trajectory of a rhodotron (TT300) \cite{cite:rhodo_design}.} 
    \label{fig:magnet_design_illustrations}
    \vspace{-15pt}
\end{figure}

>>>>

In rhodotron accelerators, magnets are used for not only guiding the electron beam to desired path, but also for synchronizing them with the RF field in the cavity. 
This syncronization, also called phase stability, is the fundamental constraint of magnet design. 
Time spent outside of the RF cavity should be precisely tuned to maximize acceleration in the succeeding pass. 
\iffalse \begin{figure}[H]
    %\captionsetup[subfigure]{justification=centering}
    %\captionsetup{justification=centering}
    \centering
    \begin{subfigure}{.5\textwidth}
      \centering
      \includegraphics[width=.9\linewidth]{../../../figures/design/Ltot.png}
    \end{subfigure}%
    \centering
    \begin{subfigure}{.5\textwidth}
      \centering
      \includegraphics[width=.9\linewidth]{../../../figures/design/tt300_mag.png}
    \end{subfigure}
    \caption{Geometry and modeling, outside trajectory of a rhodotron (TT300) \cite{rhodo_design}.}
    \label{fig:magnet_design_illustrations}
\end{figure} \fi
\vspace{20pt}
\begin{figure}[H]
    \centering
    \subfigure{{\includegraphics[width=.45\textwidth]{../../../figures/design/Ltot.png} }}%
    \qquad\subfigure{{\includegraphics[width=.45\textwidth]{../../../figures/design/tt300_mag.png} }}%
    \vspace{0pt}
    \caption{\centering Geometry and modeling, outside trajectory of a rhodotron (TT300) \cite{rhodo_design}.} 
    \label{fig:magnet_design_illustrations}
    \vspace{-15pt}
\end{figure}

### 2024-03-23 subsections/production/technical_drawings/techd.tex
Spiral corrected
hard copy ready

\begin{figure}
    \centering
    \includegraphics[angle=270,origin=c, width=.95\linewidth]{../../../figures/teknikcizim/Rho-A1.0.00.pdf}
    \vspace{20pt}
    \caption{Technical drawing of the rhodotron cavity \cite{sinan}.}
    \label{fig:techd_rhod}
\end{figure}

\begin{figure}
    \centering
    \includegraphics[angle=270,origin=c, width=.95\linewidth]{../../../figures/teknikcizim/Rhodotron Elektron Hizlandirici Sistemi.pdf}
    \vspace{20pt}
    \caption{Technical drawing of middle part of the cavity which contains the beam line \cite{sinan}.}
    \label{fig:techd_mid}
\end{figure}
%%%%%%

>>>>

\vspace{-20pt}
\begin{figure}[H]
    \centering
    \includegraphics[angle=90,origin=c, width=.95\linewidth]{../../../figures/teknikcizim/Rho-A1.0.00.pdf}
    \vspace{5pt}
    \caption{Technical drawing of the rhodotron cavity \cite{sinan}.}
    \label{fig:techd_rhod}
\end{figure}
\clearpage
\begin{figure}[!th]
    \centering
    \includegraphics[angle=90,origin=c, width=.95\linewidth]{../../../figures/teknikcizim/Rhodotron Elektron Hizlandirici Sistemi.pdf}
    \vspace{10pt}
    \caption{Technical drawing of middle part of the cavity which contains the beam line \cite{sinan}.}
    \label{fig:techd_mid}
\end{figure}

%%%%%%

### 2024-03-23 subsections/simulation/intermediate_versions/intermediate.tex
hard copy ready

\appendix
\chapter{Intermediate Codes} \label{appendix:intermediate_codes}
\begin{figure}[H]
    \centering
    %\captionsetup{justification=centering}
    \begin{subfigure}{\textwidth}
        \centering
        \begin{minted}[linenos=true, autogobble, frame=lines, framesep=2mm, fontsize=\scriptsize]{c++}
            for(double i = 2; i < 9; i += dT_out){
                t_dum += i;
                double enow = gecis(r_pos, vel, Et, t_dum);
                if( enow > maxE ){
                  maxE = enow;
                  t_opt = i;
                }
                t_dum = t;
            }
        \end{minted}
    \end{subfigure} 
    \\
    \begin{subfigure}{\textwidth}
        \centering
        \begin{minted}[linenos=true, autogobble, frame=lines, framesep=2mm,  fontsize=\scriptsize]{c++}
            double gecis(double r_pos, double vel, double Et, double &t){
                for(; r_pos >= -R2 && r_pos <= R2 ; t+=dT){
                    vel = c*sqrt(Et*Et-E0*E0)/Et;
                    double RelBeta  = vel/c;
                    double RelGamma = 1.0 / sqrt(1.0-RelBeta*RelBeta);

@@

\begin{figure}[H]
    %\captionsetup[subfigure]{justification=centering}
    %\captionsetup{justification=centering}
    \begin{subfigure}{\textwidth}
        \begin{minted}[linenos=true, autogobble, frame=lines, framesep=2mm, fontsize=\scriptsize]{c++}
            double vector3d::operator* (const vector3d& other){
                double dot = 0;
                dot += this->x * other.x;
                dot += this->y * other.y;
                dot += this->z * other.z;
                return dot;
            }
        \end{minted}
    \end{subfigure}

    \begin{subfigure}{\textwidth}
        \begin{minted}[linenos=true, autogobble, frame=lines, framesep=2mm, fontsize=\scriptsize]{c++}
            vector3d vector3d::operator% (const vector3d& other){
                double x_ = (this->y * other.z) - (this->z * other.y);
                double y_ = (this->z * other.x) - (this->x * other.z);
                double z_ = (this->x * other.y) - (this->y * other.x);
                vector3d crossed(x_, y_, z_);
                return crossed;
            }
        \end{minted}
    \end{subfigure}
    \caption{* and \% operators of vector3d class.}
    \label{fig:vector3d_dot_cross_product}
\end{figure}

\begin{figure}[H]
    \centering
    %\captionsetup{justification=centering}
    \begin{minted}[linenos=true, autogobble, frame=lines, framesep=2mm, fontsize=\scriptsize]{c++}
        bool isInsideHalfSphere(vector3d e_position, double r, vector3d hs_position){
            vector3d relative = e_position - hs_position;       
            // r/5 can be changed, use this for now                                
            if ( relative.magnitude() <= r && relative * hs_position.direction() >= -r/5){      
                return true;                                                                 
            }
            return false;
        }
    \end{minted}
\caption{Logic of is \e inside the shape of magnet.}
\label{fig:is_inside_halfsphere}
\end{figure}

\begin{figure}[H]
    \centering
    %\captionsetup{justification=centering}
    \begin{minted}[linenos=true, autogobble, frame=lines, framesep=2mm, fontsize=\scriptsize]{c++}
        vector3d Electron2D::interactB_RK(const MagneticField& B, double time_interval){
            if (B.isInside(pos) == -1){
                return vector3d(0,0,0);
            }
            Electron2D e_dummy;
            e_dummy.Et = Et;
            e_dummy.pos = pos;
            e_dummy.vel = vel;
            double time_halved = time_interval*0.5;
            // get k1                                       
            vector3d F_m = (e_dummy.vel % B.getField(pos))*eQMratio;                                       
            vector3d k1 = F_m * e_dummy.gamma_inv();     
            // get k2
            e_dummy.move(time_halved);
            e_dummy.accelerate(k1, time_halved);                                              
            F_m = (e_dummy.vel % B.getField(pos))*eQMratio;                                               
            vector3d k2 = F_m * e_dummy.gamma_inv();    
            // get k3
            e_dummy.vel = vel;
            e_dummy.accelerate(k2, time_halved);
            vector3d k3 = F_m * e_dummy.gamma_inv();   
            // get k4
            e_dummy.vel = vel;
            e_dummy.move(time_halved);
            e_dummy.accelerate(k3, time_interval);                                           
            F_m = (e_dummy.vel % B.getField(pos))*eQMratio;  
            vector3d k4 = F_m * e_dummy.gamma_inv();

            return (k1 + k2*2 + k3*2 + k4)/6;
        }
    \end{minted}
\caption{RK4-1 implemenation of \e-$\vecthreeBF{B}$.}
\label{fig:rk1_B}
\end{figure}
\begin{figure}[H]
    \centering
    %\captionsetup{justification=centering}
    \begin{minted}[linenos=true, autogobble, frame=lines, framesep=2mm, fontsize=\scriptsize]{c++}
        void Electron2D::interactRK_ActorE(const RFField& E, const MagneticField& B, double time_interval){
            vector3d run_kut_E = interactE_RK(E, time_interval);
            vector3d run_kut_B = interactB_RK(B, time_interval);

@@

\begin{figure}[H]
    \centering
    %\captionsetup{justification=centering}
    \begin{minted}[linenos=true, autogobble, frame=lines, framesep=2mm, fontsize=\tiny]{c++}
        void Gun::fireAllWithFireTimesMT(){
            std::random_device rd;
            std::mt19937 e2(rd());
            std::normal_distribution<double> Edist(Ein, sEin);

@@

\begin{figure}[H]
    \centering
    %\captionsetup{justification=centering}
    \begin{minted}[linenos=true, autogobble, frame=lines, framesep=2mm, fontsize=\tiny]{c++}
        void threadLoop(ThreadArguments thread_arguments){
            uint64_t count = 0;

@@

\begin{figure}[H]
    \centering
    %\captionsetup{justification=centering}
    \begin{minted}[linenos=true, autogobble, frame=lines, framesep=2mm, fontsize=\tiny]{c++}
        void UIThreadWork(UIThreadArgs args){
            int UI_WORK_PIECE = SIM_WORK_MASK;
            if ( !args.isService ) {
                UI_WORK_PIECE = 50;
            }

@@

\begin{figure}[H]
    \centering
    %\captionsetup{justification=centering}
    \begin{minted}[linenos=true, autogobble, frame=lines, framesep=2mm, fontsize=\scriptsize, style=staroffice]{c++}
        Optimal phase with the least RMS : -5

@@

\begin{figure}[H]
    \centering
    %\captionsetup{justification=centering}
    \begin{minted}[linenos=true, autogobble, frame=lines, framesep=2mm, fontsize=\scriptsize, style=staroffice]{c++}
        Optimal phase with the least RMS : 0

@@

\begin{figure}[H]
    \centering
    %\captionsetup{justification=centering}
    \begin{minted}[linenos=true, autogobble, frame=lines, framesep=2mm, fontsize=\tiny, style=staroffice]{c++}
        #            Rhodotron Simulation Configuration File
        # ================================================================
        #    M.Furkan Er                                     22/09/2022   
        # ================================================================
        #
        # emax = Maximum electric field strength (MV/m)
        # ein = Energy of electrons coming out of the gun (MeV)
        # einstd = Standard deviation of energy of electrons coming out of the gun (MeV)
        # targeten = Max energy on the output gif (MeV)
        # freq = Frequency of the RF field (MHz)
        # phaselag = phase lag of the first electrons (degree)
        # starttime = time to start firing the gun (ns)
        # endtime = ns to run the simulation (ns)
        # dt = time interval to do the calculations (ns)
        # guntime = how long a gun pulse is (ns)
        # gunperiod = time between two gun pulses (ns)
        # enum = number of electrons to simulate in a bunch
        # bunchnum = number of times the gun fires
        # r1 = radius of the inner cylinder (m)
        # r2 = radius of the outer cylinder (m)
        # epath = path to store the electric field data
        # bpath = path to store the magnetic field data
        # cpath = path to store the settings
        # ppath = path to store electron data
        # multh = enable or disable multitheading
        # thcount = set the maximum thread to be used
        # magrotation = degrees of rotation to enter each magnet 
        # addmagnet = takes 3 input. (B , R, < Radial distance of center >)
        # output = output file name

@@

\begin{figure}[H]
    \centering
    %\captionsetup{justification=centering}
    \begin{minted}[linenos=true, autogobble, frame=lines, framesep=2mm, fontsize=\scriptsize, style=staroffice]{c++}
        -- Simulation Configuration --
        Emax : 0.96     MV/m
        Freq : 107.5    MHz
        Phase Lag : -5  degree
        EndTime : 100   ns
        dT : 0.0001     ns
        guntime : 0.8   ns
        gunperiod : 9.3 ns
        enum : 100
        bunchnum : 2
        R1 : 0.188241   m
        R2 : 0.752967   m
        Magnet count :  5
        Ein : 0.04      MeV
        TargetE : 2.5   MeV
        --------------------------------

>>>>

\appendix
\chapter{Intermediate Codes} \label{appendix:intermediate_codes}
\begin{figure}[H]
    \centering
    %\captionsetup{justification=centering}
    \begin{subfigure}{\textwidth}
        \centering
        \begin{minted}[linenos=false, autogobble, frame=lines, framesep=2mm, fontsize=\scriptsize]{c++}
            for(double i = 2; i < 9; i += dT_out){
                t_dum += i;
                double enow = gecis(r_pos, vel, Et, t_dum);
                if( enow > maxE ){
                  maxE = enow;
                  t_opt = i;
                }
                t_dum = t;
            }
        \end{minted}
    \end{subfigure} 
    \\
    \begin{subfigure}{\textwidth}
        \centering
        \begin{minted}[linenos=false, autogobble, frame=lines, framesep=2mm,  fontsize=\scriptsize]{c++}
            double gecis(double r_pos, double vel, double Et, double &t){
                for(; r_pos >= -R2 && r_pos <= R2 ; t+=dT){
                    vel = c*sqrt(Et*Et-E0*E0)/Et;
                    double RelBeta  = vel/c;
                    double RelGamma = 1.0 / sqrt(1.0-RelBeta*RelBeta);

@@

\begin{figure}[H]
    %\captionsetup[subfigure]{justification=centering}
    %\captionsetup{justification=centering}
    \begin{subfigure}{\textwidth}
        \begin{minted}[linenos=false, autogobble, frame=lines, framesep=2mm, fontsize=\scriptsize]{c++}
            double vector3d::operator* (const vector3d& other){
                double dot = 0;
                dot += this->x * other.x;
                dot += this->y * other.y;
                dot += this->z * other.z;
                return dot;
            }
        \end{minted}
    \end{subfigure}

    \begin{subfigure}{\textwidth}
        \begin{minted}[linenos=false, autogobble, frame=lines, framesep=2mm, fontsize=\scriptsize]{c++}
            vector3d vector3d::operator% (const vector3d& other){
                double x_ = (this->y * other.z) - (this->z * other.y);
                double y_ = (this->z * other.x) - (this->x * other.z);
                double z_ = (this->x * other.y) - (this->y * other.x);
                vector3d crossed(x_, y_, z_);
                return crossed;
            }
        \end{minted}
    \end{subfigure}
    \caption{* and \% operators of vector3d class.}
    \label{fig:vector3d_dot_cross_product}
\end{figure}

\begin{figure}[H]
    \centering
    %\captionsetup{justification=centering}
    \begin{minted}[linenos=false, autogobble, frame=lines, framesep=2mm, fontsize=\scriptsize]{c++}
        bool isInsideHalfSphere(vector3d e_position, double r, vector3d hs_position){
            vector3d relative = e_position - hs_position;       
            // r/5 can be changed, use this for now                                
            if ( relative.magnitude() <= r && relative * hs_position.direction() >= -r/5){      
                return true;                                                                 
            }
            return false;
        }
    \end{minted}
\caption{Logic of is \e inside the shape of magnet.}
\label{fig:is_inside_halfsphere}
\end{figure}

\begin{figure}[H]
    \centering
    %\captionsetup{justification=centering}
    \begin{minted}[linenos=false, autogobble, frame=lines, framesep=2mm, fontsize=\scriptsize]{c++}
        vector3d Electron2D::interactB_RK(const MagneticField& B, double time_interval){
            if (B.isInside(pos) == -1){
                return vector3d(0,0,0);
            }
            Electron2D e_dummy;
            e_dummy.Et = Et;
            e_dummy.pos = pos;
            e_dummy.vel = vel;
            double time_halved = time_interval*0.5;
            // get k1                                       
            vector3d F_m = (e_dummy.vel % B.getField(pos))*eQMratio;                                       
            vector3d k1 = F_m * e_dummy.gamma_inv();     
            // get k2
            e_dummy.move(time_halved);
            e_dummy.accelerate(k1, time_halved);                                              
            F_m = (e_dummy.vel % B.getField(pos))*eQMratio;                                               
            vector3d k2 = F_m * e_dummy.gamma_inv();    
            // get k3
            e_dummy.vel = vel;
            e_dummy.accelerate(k2, time_halved);
            vector3d k3 = F_m * e_dummy.gamma_inv();   
            // get k4
            e_dummy.vel = vel;
            e_dummy.move(time_halved);
            e_dummy.accelerate(k3, time_interval);                                           
            F_m = (e_dummy.vel % B.getField(pos))*eQMratio;  
            vector3d k4 = F_m * e_dummy.gamma_inv();

            return (k1 + k2*2 + k3*2 + k4)/6;
        }
    \end{minted}
\caption{RK4-1 implemenation of \e-$\vecthreeBF{B}$.}
\label{fig:rk1_B}
\end{figure}
\begin{figure}[H]
    \centering
    %\captionsetup{justification=centering}
    \begin{minted}[linenos=false, autogobble, frame=lines, framesep=2mm, fontsize=\scriptsize]{c++}
        void Electron2D::interactRK_ActorE(const RFField& E, const MagneticField& B, double time_interval){
            vector3d run_kut_E = interactE_RK(E, time_interval);
            vector3d run_kut_B = interactB_RK(B, time_interval);

@@

\begin{figure}[H]
    \centering
    %\captionsetup{justification=centering}
    \begin{minted}[linenos=false, autogobble, frame=lines, framesep=2mm, fontsize=\tiny]{c++}
        void Gun::fireAllWithFireTimesMT(){
            std::random_device rd;
            std::mt19937 e2(rd());
            std::normal_distribution<double> Edist(Ein, sEin);

@@

\begin{figure}[H]
    \centering
    %\captionsetup{justification=centering}
    \begin{minted}[linenos=false, autogobble, frame=lines, framesep=2mm, fontsize=\tiny]{c++}
        void threadLoop(ThreadArguments thread_arguments){
            uint64_t count = 0;

@@

\begin{figure}[H]
    \centering
    %\captionsetup{justification=centering}
    \begin{minted}[linenos=false, autogobble, frame=lines, framesep=2mm, fontsize=\tiny]{c++}
        void UIThreadWork(UIThreadArgs args){
            int UI_WORK_PIECE = SIM_WORK_MASK;
            if ( !args.isService ) {
                UI_WORK_PIECE = 50;
            }

@@

\begin{figure}[H]
    \centering
    %\captionsetup{justification=centering}
    \begin{minted}[linenos=false, autogobble, frame=lines, framesep=2mm, fontsize=\scriptsize, style=staroffice]{c++}
        Optimal phase with the least RMS : -5

@@

\begin{figure}[H]
    \centering
    %\captionsetup{justification=centering}
    \begin{minted}[linenos=false, autogobble, frame=lines, framesep=2mm, fontsize=\scriptsize, style=staroffice]{c++}
        Optimal phase with the least RMS : 0

@@

\begin{figure}[H]
    \centering
    %\captionsetup{justification=centering}
    \begin{minted}[linenos=false, autogobble, frame=lines, framesep=2mm, fontsize=\tiny, style=staroffice]{c++}
        #            Rhodotron Simulation Configuration File
        # ================================================================
        #    M.Furkan Er                                     22/09/2022   
        # ================================================================
        #
        # emax = Maximum electric field strength (MV/m)
        # ein = Energy of electrons coming out of the gun (MeV)
        # einstd = Standard deviation of energy of electrons coming out of the gun (MeV)
        # targeten = Max energy on the output gif (MeV)
        # freq = Frequency of the RF field (MHz)
        # phaselag = phase lag of the first electrons (degree)
        # starttime = time to start firing the gun (ns)
        # endtime = ns to run the simulation (ns)
        # dt = time interval to do the calculations (ns)
        # guntime = how long a gun pulse is (ns)
        # gunperiod = time between two gun pulses (ns)
        # enum = number of electrons to simulate in a bunch
        # bunchnum = number of times the gun fires
        # r1 = radius of the inner cylinder (m)
        # r2 = radius of the outer cylinder (m)
        # epath = path to store the electric field data
        # bpath = path to store the magnetic field data
        # cpath = path to store the settings
        # ppath = path to store electron data
        # multh = enable or disable multitheading
        # thcount = set the maximum thread to be used
        # magrotation = degrees of rotation to enter each magnet 
        # addmagnet = takes 3 input. (B , R, < Radial distance of center >)
        # output = output file name

@@

\begin{figure}[H]
    \centering
    %\captionsetup{justification=centering}
    \begin{minted}[linenos=false, autogobble, frame=lines, framesep=2mm, fontsize=\scriptsize, style=staroffice]{c++}
        -- Simulation Configuration --
        Emax : 0.96     MV/m
        Freq : 107.5    MHz
        Phase Lag : -5  degree
        EndTime : 100   ns
        dT : 0.0001     ns
        guntime : 0.8   ns
        gunperiod : 9.3 ns
        enum : 100
        bunchnum : 2
        R1 : 0.188241   m
        R2 : 0.752967   m
        Magnet count :  5
        Ein : 0.04      MeV
        TargetE : 2.5   MeV
        --------------------------------
